\chapter{Conclusions}
\label{chap:conclusions}

My main objective was and still is to provide a solution for reducing the time required for a global, codebase-level reevaluation of static analysis after a change occurs.

The elaborated framework can transform a rather large source code repository as a whole into a graph representation and maintain it subsequently. It is found that the approach is suitable for performing code convention compliance checks and for executing static analysis tests on the graph representation.

This approach also utilizes incremental processing with file-level granularity, speeding up the static analysis. Based on my measurements the framework is fast enough to help its users with fast changing code repositories.

Once the framework contains enough transformations and queries for handling the language, it can extend the everyday toolkit of developers.

\section{Summary of Contributions}
I presented an extensible proof of my novel concept to perform incremental static analysis on dynamic JavaScript source code repositories based on graph transformations. My proposal is based on software code modeling, graph transformation and graph pattern matching. The feasibility of the approach was evaluated using manual validation and benchmarking.

\subsection{Scientific Contributions}
I have achieved the following scientific contributions:

\begin{itemize}[topsep=0pt]
	\item Proposed an architecture for building an incremental static analyzer using freely available components.
	\item Proposed an approach to transform JavaScript source code repository into a connected graph model.
	\item Provided an algorithm to update the graph data model incrementally.
	\item Presented a working method for finding problematic code parts using graph pattern matching.
	\item Provided a transformation approach for creating CFG\footnote{Control Flow Graph} from ASG\footnote{Abstract Semantic Graph}.
\end{itemize}

\subsection{Practical Accomplishments}
I have also achieved the following practical accomplishments:

\begin{itemize}[topsep=0pt]
	\item Created an extensible incremental static analysis framework.
	\item Developed a tool for transforming JavaScript source code into graph data model.
	\item Designed a benchmark evaluating the approach.
	\item The prototype framework has been published as an open-source project\footnote{\small\url{https://github.com/ftsrg/codemodel-rifle}}.
\end{itemize}

\section{Novel Results}
Compared to my earlier work in static analysis~\cite{stein-daniel-tdk,stein-daniel-bsc,stein-daniel-ttc}, this thesis has the following novel contributions:

\begin{itemize}[topsep=0pt]
	\item This approach utilizes a property graph for storing the transformed model of the source code (as opposed to EMF\footnote{Eclipse Modeling Framework} and RDF\footnote{Resource Description Framework} models).

	\item My approach tackles the problem of carrying out static analysis for dynamic languages (e.g., JavaScript), instead of static languages (e.g., Java).

	\item Besides static analysis, I proposed an algorithm for transforming the ASG into a variant of the CFG.

	\item I investigated the feasibility of a type inferring system on the graph structure using graph transformation.

	\item This approach was tested on an industrial case study, provided by Tresorit Kft.

	\item I implemented a proof-of concept IDE integration with Visual Studio Code that is able to detect dead code with quick response times and display corresponding warnings to the source code developers.
\end{itemize}
